---
title: "COVID Support Loss Analytic Code"
output:
  html_document: default
---

```{r, include = F}
####  Setup  ####
library(easypackages)
libraries("tidyverse", "ggplot2", "readr", "magrittr", "Amelia", "scales", "questionr", "corrplot", "knitr", "kableExtra", "conflicted", "fastDummies")
walk(c("filter", "select", "count", "group_by"), ~ conflict_prefer(., "dplyr"))

rm(list = ls())

opts_chunk$set(echo = F, message = F)
opts <- options(knitr.kable.NA = "")



####  Load Data  ####
raw <- read_rds("G:\\Shared drives\\Psychology_JSLab\\Projects\\In_Progress\\SBU_COVID19_Study\\Clean Data\\SBU COVID-19 Survey - Baseline.rds")



####  Clean Data  ####
clean <- raw %>%
  
  ## Filter TBD
  filter(
    
    position == 1, #Just undergraduate students
    supportLoss %in% 1:3 #Drop respondents who either didn't respond to support loss question, or don't usually receive support
    
    
  ) %>%
  
  ## Recode and rename variables as necessary and select these variables for analysis
  mutate(
    
    supportLoss = supportLoss %in% 1:2,
    sex = case_when(sex == 1 ~ "Male",
                    sex == 2 ~ "Female",
                    T ~ NA_character_),
    ethnicity = case_when(race == 1 & hispanic == 0 ~ "White non-Hispanic",
                          race == 2 & hispanic == 0 ~ "Black non-Hispanic",
                          race == 3 & hispanic == 0 ~ "AI/AN non-Hispanic",
                          race == 4 & hispanic == 0 ~ "Asian non-Hispanic",
                          race == 5 & hispanic == 0 ~ "NH/PI non-Hispanic",
                          race == 6 & hispanic == 0 ~ "Multi non-Hispanic",
                          hispanic == 1 ~ "Hispanic",
                          T ~ NA_character_),
    genderMinority = {gender == 3 | (genderDiff == 1 & .$sex != gender)} %>%
      if_else(is.na(gender), F, .),
    sexualMinority = if_else(!is.na(sexualOrientation), sexualOrientation %in% 2:9, NA),
    GAD7 = GAD_1 + GAD_2 + GAD_3 + GAD_4 + GAD_5 + GAD_6 + GAD_7 - 7, #-7 to account for original coding being off by 1
    PHQ = PHQ_1 + PHQ_2 + PHQ_3 + PHQ_4 + PHQ_5 + PHQ_6 + PHQ_7 + PHQ_8 + PHQ_9 - 9, #-9 to account for original coding being off by 1
    SI = PHQ_9 > 1 #Any response greater than "Not at all," which is coded as a 1, not a 0, in the raw data
  
  ) %>%
  
  select_at(vars(supportLoss, sex, ethnicity, genderMinority, sexualMinority, GAD7, PHQ, SI,
                 matches("^GAD_[0-9]"),
                 matches("^PHQ_[0-9]"),
                 matches("^impact.")))

before <- nrow(clean)

df <- clean[complete.cases(clean),]

after <- nrow(df)
```

# Study Overview

This study investigates loss of mental, behavioral, and/or emotional support in the midst of the COVID-19 pandemic among undergraduate students at a large university. The goal of this study is to gather evidence on the extent of this loss and whether it differs by individual characteristics. Specifically, we seek to answer the following questions:

1. What is the extent of loss of support during the COVID-19 pandemic among members of a large university community?
2. Does loss of support differ by (a) sociodemographic factors or (b) levels of depression and anxiety symptomatology, including suicidal ideation?

This is an exploratory study and we do not have specific hypotheses regarding the extent of support loss, or whether/how it differs across the above conditions.

Analyses are limited to complete cases only. This drops `r before - after` of `r before` cases.

# Sample Description

Categorical variables (report n, percent)

```{r}
df %>%
  select(-matches("GAD"),
         -matches("PHQ"),
         -matches("impact")) %>%
  mutate_all(as.character) %>%
  pivot_longer(everything()) %>%
  count(name, value) %>%
  group_by(name) %>%
  mutate(pct = percent(n / sum(n), accuracy = .01))
```

Continuous variables (report mean, sd)

```{r}
df %>%
  select(GAD7, PHQ) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  summarize(mean = mean(value),
            sd = sd(value),
            `>= 10` = mean(value >= 10),
            `>= 15` = mean(value >= 15),
            `>= 20` = mean(value >= 20))
```

# Support loss by COVID impacts

```{r}
df %>%
  select_at(vars(supportLoss, matches("^impact."))) %>%
  pivot_longer(matches("^impact."), names_to = "impactType", values_to = "impactPresence") %>%
  group_by(impactType) %>%
  summarize(impactPresence.m = mean(impactPresence),
            impactPresence.n = sum(impactPresence),
            n = n(),
            impactPresence = paste0(percent(impactPresence.m, .01), " (", impactPresence.n, " of ", n, ")")) 

df %>%
  select_at(vars(supportLoss, matches("^impact."))) %>%
  pivot_longer(matches("^impact."), names_to = "impactType", values_to = "impactPresence") %>%
  group_by(supportLoss, impactType) %>%
  summarize(impactPresence.m = mean(impactPresence),
            impactPresence.n = sum(impactPresence),
            n = n(),
            impactPresence = paste0(percent(impactPresence.m, .01), " (", impactPresence.n, " of ", n, ")")) %>%
  pivot_wider(id_cols = impactType, names_from = supportLoss, values_from = impactPresence) %>%
  mutate(impactType = gsub("impact.", "", impactType)) %>%
  select(`Impact Type` = impactType,
         `Support Loss` = `TRUE`,
         `No Loss` = `FALSE`)
```

# Internal consistency

```{r}
df %>%
  select(GAD_1:GAD_7) %>%
  psych::alpha() %>%
  .[["total"]] %>%
  kable(digits = 2, caption = "Internal consistency for GAD7") %>%
  kable_styling()

df %>%
  select(PHQ_1:PHQ_9) %>%
  psych::alpha() %>%
  .[["total"]] %>%
  kable(digits = 2, caption = "Internal consistency for PHQ") %>%
  kable_styling()
```

# Main analysis

```{r}
printLoss <- function(x) {
  
  pct = percent(mean(x), accuracy = .1)
  n = length(x)
  
  out = paste0(pct, " (n=", n, ")")
  
  return(out)
  
}

out <- tibble_row(
  
  group = "Full sample",
  lostSupport = printLoss(df$supportLoss),
  statistic = NA_character_,
  p = NA_real_
  
)

vars = c("sex", "ethnicity", "genderMinority", "sexualMinority", "GAD7", "PHQ", "SI")

for(var in vars) {
  
  class = class(df[[var]])
  
  if(class %in% c("character", "logical")) {
    
    table <- table(df$supportLoss, df[[var]])
    test <- chisq.test(table)
    
    print(paste0("Expected values for the Chi-square test of supportLoss, ", var))
    print(test$expected)
    
    statistic <- paste0(names(test$statistic),
                        "(df=", test$parameter[[1]], ") = ", 
                        round(test$statistic[[1]], 3))
    p <- test$p.value[[1]]
    
    out <- rbind(out, tibble(
      group = var,
      lostSupport = NA_character_,
      statistic = statistic,
      p = p
      
    ))
    
    for(level in unique(df[[var]])) {
      
      out <- rbind(out, tibble(
        group = level,
        lostSupport = printLoss(df$supportLoss[df[var] == level]),
        statistic = NA_character_,
        p = NA_real_
        
      ))
      
    }
    
  } else {
    
    x <- df[[var]][df$supportLoss == 1]
    y <- df[[var]][df$supportLoss == 0]
    
    test <- t.test(x, y,
                   alternative = "two.sided",
                   paired = F,
                   var.equal = F)
    
    statistic <- paste0(names(test$statistic), 
                        "(df=", round(test$parameter[[1]], 1), ") = ",
                        round(test$statistic[[1]], 3))
    p = test$p.value
    
        out <- rbind(out, tibble(
      group = var,
      lostSupport = NA_character_,
      statistic = statistic,
      p = p
      
    ))
    
  }

}

out$p[!is.na(out$p)] <- p.adjust(out$p[!is.na(out$p)], "BH") 
out$p[!is.na(out$p)] <- paste0("p = ", round(out$p[!is.na(out$p)], 3))

out %>%
  kable() %>%
  kable_styling()
```

A quick table to aid in the interpretation of above t-tests

```{r}
df %>%
  group_by(supportLoss) %>%
  summarize(n = n(),
            GAD7 = paste0(round(mean(GAD7), 2), " (",
                          round(sd(GAD7), 2), ")"),
            PHQ = paste0(round(mean(PHQ), 2), " (",
                         round(sd(PHQ), 2), ")")) %>%
  kable(digits = 2) %>%
  kable_styling()
```

And a plot to this effect

```{r}
df %>%
  select(`Support Loss` = supportLoss,
         `GAD-7` = GAD7, 
         `PHQ-9` = PHQ, 
         SI) %>%
  pivot_longer(c(`GAD-7`, `PHQ-9`, SI)) %>%
  group_by(`Support Loss`, name) %>%
  summarize(n = n(),
            Mean = mean(value),
            se = sd(value) / sqrt(n),
            .groups = "drop") %>%
  ggplot() +
    geom_point(aes(`Support Loss`, Mean)) +
    geom_errorbar(aes(x = `Support Loss`, ymin = Mean - se, ymax = Mean + se)) +
    facet_wrap(~ name, scales = "free_y") +
    scale_y_continuous(name = NULL, limits = c(0, NA)) +
    theme_classic() +
    ggtitle("Mental Health Symptoms by Support Loss", 
            subtitle = "Means and Standard Errors")
```

Another pair of tables on the SI - support loss relationship

```{r}
df %>%
  group_by(SI) %>%
  count(supportLoss) %>%
  mutate(pct = percent(n / sum(n)))

df %>%
  group_by(supportLoss) %>%
  count(SI) %>%
  mutate(pct = percent(n / sum(n)))
```